\textsf{Are large language models superhuman chemists?}

### 2024-10-24 src/tex/appendix.tex
suggestions from gil

\begin{figure}[htb]
    \centering
    \includegraphics{figures/overall_performance.pdf}
    \caption{\textbf{Overall performance of the models on the \chembench corpus.} The bar plot shows the fraction of questions that were answered completely correctly by the models. Scores computed on the entire \chembench corpus.}
    \label{fig:barplot_all_correct_all_questions}
    \script{plot_overview_performance_plot.py}
\end{figure}

@@

\begin{table}
    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
    \resizebox{\textwidth}{!}{
    \variable{output/performance_table.tex}
    }
    \label{tab:performance_table}
\end{table}

\begin{table}
    \caption{\textbf{Performance of the models on the \enquote{tiny} subset of the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
    \resizebox{\textwidth}{!}{
    \variable{output/performance_table_human_subset.tex}
    }
    \label{tab:performance_table_human_subset}
\end{table}

\begin{figure}[htb]
    \centering
    \includegraphics[width=\textwidth]{figures/all_questions_models_completely_correct_radar_overall.pdf}
    \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The radar plot shows the performance of the models on the different topics of the \chembench corpus. The performance is measured as the fraction of questions answered completely correctly by the models.
    A score of 1 indicates that all questions were answered completely correctly, while a score of 0 indicates that none were answered completely correctly.
    }
    \label{fig:all_questions_models_completely_correct_radar_overall}
    \script{analyze_model_reports.py}
\end{figure}

@@

\begin{figure}[htb]
    \centering
    \includegraphics{figures/performance_per_topic.pdf}
    \caption{\textbf{Fraction of correctly answered questions per data source.} The heatmap shows, in color, the fraction of questions answered correctly by different systems for some of our data sources. The performance is measured as the fraction of questions answered completely correctly by the models. A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none of the questions were answered completely correctly.
        We see that the performance of the models varies significantly between the different data sources. For instance, it is interesting to observe that questions sourced based on textbooks seem easier for our leading models than for humans. However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
    }
    \label{fig:performance_per_topic}
    \script{analyze_performance_per_source.py}
\end{figure}

@@

\begin{figure}[htb]
    \centering
    \includegraphics{figures/performance_per_topic_tiny.pdf}
    \caption{\textbf{Fraction of completely correctly answered questions per data source on the \enquote{tiny} subset.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources. The performance is measured as the fraction of questions answered completely correctly by the models. A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none were answered completely correctly.
        We see that the performance of the models varies significantly between the different data sources. For instance, it is interesting to observe that questions sourced based on textbooks seem easier for the leading models than for humans. However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
    }
    \label{fig:performance_per_topic_tiny}
    \script{analyze_performance_per_source.py}
\end{figure}

@@

\clearpage
\subsection{Influence of model scale}
\Cref{fig:model_size_plot} shows the performance of the models as a function of the number of parameters in the model.
We see that the performance of the models correlates with their size for the models of the LLama-3 and Llama-3.1 herd of models.

@@

In the applications, human participants were presented with molecules as rendered drawings and SMILES strings. \LaTeX\xspace equations and chemical equations were rendered using MathJax (\Cref{fig:screenshots}).

@@

\begin{figure}[htb]
    \centering
    \includegraphics{figures/human_score_distribution.pdf}
    \script{plot_human_score_distribution.py}
    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered completely correctly.
    Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.}
    \label{fig:human_score_distribution}
\end{figure}

We also recorded the time humans took to answer the questions. This time is the time from the question being displayed to the human to the human submitting the answer.

@@

\clearpage
\subsection{Leaderboard}
Our leaderboard is based on the tool chain developed for Matbench.\autocite{Dunn_2020}
Briefly, the \chembench pipeline produces standardized files in \texttt{json} format that contributors can add via pull requests to the \chembench repository.
The Markdown tables and interactive plots are automatically generated and updated on the \chembench website. The leaderboard is available at \url{https://lamalab-org.github.io/chem-bench/leaderboard/}.

>>>>

\begin{figure}[htb]
    \centering
    \includegraphics{figures/overall_performance.pdf}
    \caption{\textbf{Overall performance of the models on the \chembench corpus.} The bar plot shows the fraction of questions that were answered correctly by the models. Scores computed on the entire \chembench corpus.}
    \label{fig:barplot_all_correct_all_questions}
    \script{plot_overview_performance_plot.py}
\end{figure}

@@

\begin{table}
    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.}
    \resizebox{\textwidth}{!}{
    \variable{output/performance_table.tex}
    }
    \label{tab:performance_table}
\end{table}

\begin{table}
    \caption{\textbf{Performance of the models on the \enquote{tiny} subset of the \chembench corpus.} The table shows the fraction of questions answered correctly by the models for different skills and difficulty levels.}
    \resizebox{\textwidth}{!}{
    \variable{output/performance_table_human_subset.tex}
    }
    \label{tab:performance_table_human_subset}
\end{table}

\begin{figure}[htb]
    \centering
    \includegraphics[width=\textwidth]{figures/all_questions_models_completely_correct_radar_overall.pdf}
    \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The radar plot shows the performance of the models on the different topics of the \chembench corpus. The performance is measured as the fraction of questions answered correctly by the models.
    A score of 1 indicates that all questions were answered correctly, while a score of 0 indicates that none were answered correctly.
    }
    \label{fig:all_questions_models_completely_correct_radar_overall}
    \script{analyze_model_reports.py}
\end{figure}

@@

\begin{figure}[htb]
    \centering
    \includegraphics{figures/performance_per_topic.pdf}
    \caption{\textbf{Fraction of correctly answered questions per data source.} The heatmap shows, in color, the fraction of questions answered correctly by different systems for some of our data sources. The performance is measured as the fraction of questions answered correctly by the models. A score of one (red) indicates that all questions were answered correctly, while a score of zero (blue) indicates that none of the questions were answered correctly.
        We see that the performance of the models varies significantly between the different data sources. For instance, it is interesting to observe that questions sourced based on textbooks seem easier for our leading models than for humans. However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
    }
    \label{fig:performance_per_topic}
    \script{analyze_performance_per_source.py}
\end{figure}

@@

\begin{figure}[htb]
    \centering
    \includegraphics{figures/performance_per_topic_tiny.pdf}
    \caption{\textbf{Fraction of completely correctly answered questions per data source on the \enquote{tiny} subset.} The heatmap shows, in color, the fraction of questions answered correctly by different systems for some of our data sources. The performance is measured as the fraction of questions answered correctly by the models. A score of one (red) indicates that all questions were answered correctly, while a score of zero (blue) indicates that none were answered correctly.
        We see that the performance of the models varies significantly between the different data sources. For instance, it is interesting to observe that questions sourced based on textbooks seem easier for the leading models than for humans. However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
    }
    \label{fig:performance_per_topic_tiny}
    \script{analyze_performance_per_source.py}
\end{figure}

@@

\clearpage
\subsection{Influence of model scale}
\Cref{fig:model_size_plot} shows the performance of the models as a function of the number of parameters in the model.
We see that the performance of the models correlates with scale for the models of the LLama-3 and Llama-3.1 herd of models.

@@

In the applications, human participants were presented with molecules as rendered drawings and SMILES strings. \LaTeX\xspace equations and chemical equations were rendered using MathJax \Cref{fig:screenshots}.

@@

\begin{figure}[htb]
    \centering
    \includegraphics{figures/human_score_distribution.pdf}
    \script{plot_human_score_distribution.py}
    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered correctly.
    Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.}
    \label{fig:human_score_distribution}
\end{figure}

We also recorded the time humans took to answer the questions \cref{fig:human_timing}. This time is the time from the question being displayed to the human to the human submitting the answer.

@@

\clearpage
\subsection{Leaderboard}
\label{sec:leaderboard}
Our leaderboard is based on the tool chain developed for Matbench.\autocite{Dunn_2020}
Briefly, the \chembench pipeline produces standardized files in \texttt{json} format that contributors can add via pull requests to the \chembench repository.
The Markdown tables and interactive plots are automatically generated and updated on the \chembench website. The leaderboard is available at \url{https://lamalab-org.github.io/chem-bench/leaderboard/}.

### 2024-10-24 src/tex/ms.tex
suggestions from gil

We curated more than 2,800 question-answer pairs, evaluated leading open- and closed-source LLMs, and found that the best models outperformed the best human chemists (volunteers) in our study on average.
However, the models struggle with some tasks that are easy for human experts and provide overconfident predictions.

@@

However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\autocite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards, urbina2022teachable}
To some extent, this is not surprising as any technology that, for instance, is used to design non-toxic molecules can also be used inversely to predict toxic ones (even though the synthesis would still require access to controlled physical resources and facilities).
Still, it is essential to realize that such models' user base is wider than that of chemistry and materials science experts who can critically reflect on every output such models produce.
For example, many students frequently consult these tools---perhaps even to prepare chemical experiments.\autocite{Intelligent.com_2023}
This also applies to users from the general public, who might consider using \glspl{llm} to answer questions about the safety of chemicals.
Thus, for some users, misleading information---especially about safety-related aspects---might lead to harmful outcomes.
However, even for experts, chemical knowledge and reasoning capabilities are essential as they will determine the capabilities and limitations of their models in their work, e.g., in copilot systems for chemists.
Unfortunately, apart from exploratory reports such as by prompting leading models with various scientific questions,\autocite{ai4science2023impact} there is little systematic evidence on how \glspl{llm} perform compared to expert (human volunteers) chemists.

@@

While some benchmarks based on university entrance exams\autocite{Zaki_2024, arora2023llms} or automatic text mining\autocite{song2023honeybee, wei2021chemistryqa, song-etal-2023-matsci} have been proposed, none of them have been widely accepted.
This is likely because they cannot automatically be used with black box (or tool-augmented) systems, do not cover a wide range of topics and skills, or are not carefully validated by experts.
On top of that, the existing benchmarks are not designed to be used with models that support special treatment of molecules or equations and do not provide insights on how the models compare relative to experts.

@@

To contextualize the scores, we also surveyed \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human) chemists of different specializations. In parts of the survey, the volunteers were also allowed to use tools such as web search to create a realistic setting.

@@

\begin{figure}
    \includegraphics[width=\textwidth]{figures/overview_figure.pdf}
    \caption{\textbf{Overview of the \chembench framework.} The figure shows the different components of the \chembench framework.
    The framework's foundation is the benchmark corpus that consists of many questions and answers that we manually or semi-automatically compiled from various sources.
    Questions are classified based on topics, required skills (reasoning, calculation, knowledge, intuition), and difficulty levels.
    We then used this corpus to evaluate the performance of various models and tool-augmented systems using a custom framework. To provide a baseline, we built a web application that we used to survey experts in chemistry.
    The results of the evaluations are then compiled in publicly accessible leaderboards, which we propose as a foundation for evaluating future models.
    }
    \label{fig:overview_figure}
\end{figure}

@@

\paragraph{\enquote{Tiny} subset}
It is important to note that a smaller subset of the corpus might be more practical for routine evaluations.\autocite{polo2024tinybenchmarks}
For instance,~\textcite{liang2023holistic} report costs of more than \$10,000 for \gls{api} calls for a single evaluation on the widely used \gls{helm} benchmark.
To address this, we also provide a subset (\variable{output/num_human_answered_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus. While it is impossible to comprehensively represent the full corpus in a subset, we aimed to include a maximally diverse set of questions and a more balanced distribution of topics and skills (see \Cref{sec:subset-selection} for details on the curation process).
Our human volunteers answered all the questions in this subset.

@@

\paragraph{Overall system performance}
To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of leading models\autocite{Huggingface} on the \chembench corpus, including systems augmented with external tools.
An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar} (all results can be found in \Cref{tab:performance_table} and \Cref{tab:performance_table_human_subset}).
In this figure, we show the percentage of questions that the models answered correctly.
Moreover, we show the worst, best, and average performance of the experts (human volunteers) in our study, which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
Remarkably, the figure shows that the leading \gls{llm}, \oone, outperforms the best human in our study in this overall metric by almost a factor of two.
Many other models also outperform the average human performance.
Interestingly, \LlamaThreeOneFourZeroFiveBInstruct shows performance that is close the leading proprietary models, indicating that new open source models can be competitive with the best proprietary models also in chemical settings.

Notably, we find that models are still limited in their ability to answer knowledge-intensive questions (\Cref{tab:performance_table_human_subset}), that is they did not memorize the relevant facts. Our results indicate that this is not a limitation that could be overcome by simple application of \gls{rag} systems such as \PaperQATwo. This is likely because the required knowledge cannot easily be accessed via papers (which is the only external knowledge \PaperQATwo has access to) but rather by lookup in specialized databases (e.g., PubChem, Gestis), which also the humans in our study used to answer such questions (\Cref{fig:tool_use}).
This indicates that there is still room for improvement of chemical \gls{llm} by training them on more specialized data sources or by integrating them with specialized databases.

In addition, our analysis shows that the performance of models is correlated with their size (see \Cref{fig:model_size_plot}). This is in line with observations in other domains but also indicates that chemical \gls{llm} could, to some extent, be further improved by scaling them up.

\paragraph{Performance per topic} To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}) and  classified all questions in the \chembench corpus into these topics.
We then computed the percentage of questions the models or experts answered correctly for each topic and show them in \Cref{fig:all_questions_models_completely_correct_radar_human}.
In this spider chart, the worst score for every dimension is zero (no question answered correctly), and the best score is one (all questions answered correctly). Thus, a larger colored area indicates a better performance.

\begin{figure}[!h]
    \centering
    \includegraphics[width=\textwidth]{figures/all_questions_models_completely_correct_radar_human.pdf}
    \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus. The performance is measured as the fraction of questions that were answered completely correctly by the models.
    The best score for every dimension is one (all questions answered correctly), and the worst is zero (no question answered correctly). A larger colored area indicates a better performance.
    This figure shows the performance on the subset of questions that were answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
    }
    \label{fig:all_questions_models_completely_correct_radar_human}
    \script{analyze_model_reports.py}
\end{figure}

@@

 In \Cref{fig:confidence_vs_performance}, we show that for some models, there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
 For applications in which humans might rely on the models to provide answers with trustworthy uncertainty estimates, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the model's outputs.\autocite{Li_2023, miret2024llms}
 For example, for the questions about the safety profile of compounds, \GPTFour reported an average confidence of \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
 While, on average, the verbalized confidence estimates from Claude 3.5 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading in some cases.
 For example, for the questions about \gls{ghs} pictograms Claude 3.5 returns an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.

\section{Conclusions}
On the one hand, our findings underline the impressive capabilities of \glspl{llm} in the chemical sciences: Leading models outperform domain experts in specific chemistry questions on many topics.
On the other hand, there are still striking limitations.
For very relevant topics, the answers that models provide are wrong.
On top of that, many models are not able to reliably estimate their own limitations.
Yet, the success of the models in our evaluations perhaps also reveals more about the limitations of the questions we use to evaluate models---and chemists---than about the models themselves.
For instance, while models perform well on many textbook questions, they struggle with questions that require some more reasoning about chemical structures (e.g., number of isomers or \gls{nmr} peaks).
Given that the models outperformed the average human in our study, we need to rethink how we teach and examine chemistry.
Critical reasoning is increasingly essential, and rote solving of problems or memorization of facts is a domain in which \glspl{llm} will continue to outperform humans (when trained on the right training corpus).

@@

\paragraph{Study design}
Human volunteers were asked the questions in a custom-built web interface (see \Cref{sec:human_baseline}) which rendered chemicals and equations. Questions were shown in random order and volunteers were not allowed to skip questions. For a subset of the questions, the volunteers were allowed to use external tools (excluding other \gls{llm} or asking other people) to answer the questions. Prior to answering questions, volunteers were asked to provide information about their education and experience in chemistry. The study was conducted in English.

\paragraph{Volunteers}
Users were open to reporting about their experience in chemistry.
Overall, \variable{output/num_users_with_education_info.txt} did so.
Out of those,
\variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree (and are currently enrolled in Ph.D.\ studies), and \variable{output/num_human_bachelor.txt} have a bachelor's degree. For the analysis, we excluded volunteers with less than two years of experience in chemistry after their first university-level course in chemistry.

@@

\section*{Statement of ethical compliance}
The authors confirm to have complied with all relevant ethical regulations, according to the Ethics Commission of the Friedrich Schiller University Jena (which decided that study is ethically safe). Informed consent was obtained from all volunteers.

>>>>

We curated more than 2,800 question-answer pairs, evaluated leading open- and closed-source LLMs, and found that the best models outperformed the best human chemists in our study on average.
However, the models struggle with some tasks that are easy for human experts and provide overconfident predictions.

@@

However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\autocite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards, urbina2022teachable}
To some extent, this is not surprising as any technology that, for instance, is used to design non-toxic molecules can also be used inversely to predict toxic ones (even though the synthesis would still require access to controlled physical resources and facilities).
Still, it is essential to realize that LLMs' user base is wider than that of chemistry and materials science experts who can critically reflect on every output these models produce.
For example, many students frequently consult these tools---perhaps even to prepare chemical experiments.\autocite{Intelligent.com_2023}
This also applies to users from the general public, who might consider using \glspl{llm} to answer questions about the safety of chemicals.
Thus, for some users, misleading information---especially about safety-related aspects---might lead to harmful outcomes.
However, even for experts, chemical knowledge and reasoning capabilities are essential as they will determine the capabilities and limitations of their models in their work, e.g., in copilot systems for chemists.
Unfortunately, apart from exploratory reports such as by prompting leading models with various scientific questions,\autocite{ai4science2023impact} there is little systematic evidence on how \glspl{llm} perform compared to expert (human) chemists.

@@

While some benchmarks based on university entrance exams\autocite{Zaki_2024, arora2023llms} or automatic text mining\autocite{song2023honeybee, wei2021chemistryqa, song-etal-2023-matsci} have been proposed, none of them have been widely accepted.
This is likely because they cannot automatically be used with black box (or tool-augmented) systems, do not cover a wide range of topics and skills, or are not carefully validated by experts.
On top of that, the existing benchmarks are not designed to be used with models that support special treatment of molecules or equations and do not provide insights on how the models compare relative to experts \autocite{wu2018moleculenet}.

@@

To contextualize the scores, we also surveyed \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human) chemists of different specializations. In parts of the survey, the participants were also allowed to use tools such as web search to create a realistic setting.
Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but, in many cases, including safety-related ones, might be very misleading.
We find that there are still limitations for current models to overcome, such that they can be directly applied in autonomous systems for chemists.

@@

\begin{figure}
    \includegraphics[width=\textwidth]{figures/overview_figure.pdf}
    \caption{\textbf{Overview of the \chembench framework.} The figure shows the different components of the \chembench framework.
    The framework's foundation is the benchmark corpus that consists of many questions and answers that we manually or semi-automatically compiled from various sources.
    Questions are classified based on topics, required skills (reasoning, calculation, knowledge, intuition), and difficulty levels.
    We then used this corpus to evaluate the performance of various models and tool-augmented systems using a custom framework. To provide a baseline, we built a web application that we used to survey experts in chemistry.
    The results of the evaluations are then compiled in publicly accessible leaderboards \ref{sec:leaderboard}, which we propose as a foundation for evaluating future models.
    }
    \label{fig:overview_figure}
\end{figure}

@@

\paragraph{\enquote{Tiny} subset}
It is important to note that a smaller subset of the corpus might be more practical for routine evaluations.\autocite{polo2024tinybenchmarks}
For instance,~\textcite{liang2023holistic} report costs of more than \$10,000 for \gls{api} calls for a single evaluation on the widely used \gls{helm} benchmark.
To address this, we also provide a subset (\variable{output/num_human_answered_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus. While it is impossible to comprehensively represent the full corpus in a subset, we aimed to include a maximally diverse set of questions and a more balanced distribution of topics and skills (see \Cref{sec:subset-selection} for details on the curation process).
Our human participants answered all the questions in this subset.

@@

\paragraph{Overall system performance}
To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of leading models\autocite{Huggingface} on the \chembench corpus, including systems augmented with external tools.
An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar} (all results can be found in \Cref{tab:performance_table} and \Cref{tab:performance_table_human_subset}).
In this figure, we show the percentage of questions that the models answered correctly.
Moreover, we show the worst, best, and average performance of the human experts in our study, which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
Remarkably, the figure shows that the leading \gls{llm}, \oone, outperforms the best human in our study in this overall metric by almost a factor of two.
Many other models also outperform the average human performance.
Interestingly, \LlamaThreeOneFourZeroFiveBInstruct shows performance that is close the leading proprietary models, indicating that new open source models can be competitive with the best proprietary models also in chemical settings.

Notably, we find that models are still limited in their ability to answer knowledge-intensive questions (\Cref{tab:performance_table_human_subset}), that is they did not memorize the relevant facts. Our results indicate that this is not a limitation that could be overcome by simple application of \gls{rag} systems such as \PaperQATwo. This is likely because the required knowledge cannot easily be accessed via papers (which is the only external knowledge \PaperQATwo has access to) but rather by lookup in specialized databases (e.g., PubChem, Gestis), which also the humans in our study used to answer such questions (\Cref{fig:tool_use}).
This indicates that there is still room for improvement in chemical \gls{llm}s by training them on more specialized data sources or by integrating them with specialized databases.

In addition, our analysis shows that the performance of models is correlated with their size (see \Cref{fig:model_size_plot}). This is in line with observations in other domains but also indicates that chemical \gls{llm}s could, to some extent, be further improved by scaling them up.

\paragraph{Performance per topic} To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}) and  classified all questions in the \chembench corpus into these topics.
We then computed the percentage of questions the models or humans answered correctly for each topic and show them in \Cref{fig:all_questions_models_completely_correct_radar_human}.
In this spider chart, the worst score for every dimension is zero (no question answered correctly), and the best score is one (all questions answered correctly). Thus, a larger colored area indicates a better performance.

\begin{figure}[!h]
    \centering
    \includegraphics[width=\textwidth]{figures/all_questions_models_completely_correct_radar_human.pdf}
    \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus. The performance is measured as the fraction of questions that were answered correctly by the models.
    The best score for every dimension is one (all questions answered correctly), and the worst is zero (no question answered correctly). A larger colored area indicates a better performance.
    This figure shows the performance on the subset of questions that were answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:barplot_all_correct_all_questions}.
    }
    \label{fig:all_questions_models_completely_correct_radar_human}
    \script{analyze_model_reports.py}
\end{figure}

@@

 In \Cref{fig:confidence_vs_performance}, we show that for some models, there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
 For applications in which humans might rely on the models to provide answers with trustworthy uncertainty estimates, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the model's outputs.\autocite{Li_2023, miret2024llms}
 For example, for the questions about the safety profile of compounds, \GPTFour reported an average confidence of \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
 While, on average, the verbalized confidence estimates from Claude 3 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading in some cases.
 For example, for the questions about \gls{ghs} pictograms Claude 3 returns an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.

\section{Conclusions}
On the one hand, our findings underline the impressive capabilities of \glspl{llm} in the chemical sciences: Leading models outperform domain experts in specific chemistry questions on many topics.
On the other hand, there are still striking limitations.
For very relevant topics the answers models provide are wrong.
On top of that, many models are not able to reliably estimate their own limitations.
Yet, the success of the models in our evaluations perhaps also reveals more about the limitations of the questions we use to evaluate models---and chemists---than about the models themselves.
For instance, while models perform well on many textbook questions, they struggle with questions that require some more reasoning about chemical structures (e.g., number of isomers or \gls{nmr} peaks).
Given that the models outperformed the average human in our study, we need to rethink how we teach and examine chemistry.
Critical reasoning is increasingly essential, and rote solving of problems or memorization of facts is a domain in which \glspl{llm} will continue to outperform humans (when trained on the right training corpus).

@@

\paragraph{Study design}
Human participants were asked the questions in a custom-built web interface (see \Cref{sec:human_baseline}) which rendered chemicals and equations. Questions were shown in random order and participants were not allowed to skip questions. For a subset of the questions, the participants were allowed to use external tools (excluding other \gls{llm} or asking other people) to answer the questions. Prior to answering questions, participants were asked to provide information about their education and experience in chemistry. The study was conducted in English.

\paragraph{Participants}
Users were open to reporting about their experience in chemistry.
Overall, \variable{output/num_users_with_education_info.txt} did so.
Out of those,
\variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree (and are currently enrolled in Ph.D.\ studies), and \variable{output/num_human_bachelor.txt} have a bachelor's degree. For the analysis, we excluded participants with less than two years of experience in chemistry after their first university-level course in chemistry.

@@

\section*{Statement of ethical compliance}
The authors confirm to have complied with all relevant ethical regulations, according to the Ethics Commission of the Friedrich Schiller University Jena (which decided that study is ethically safe). Informed consent was obtained from all participants.
